---
title: "Getting started"
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


[**Here, we point you to useful resources for various topics**]{style="color:red;"}

# Before You Start

*(from Justin C.)*

For this short course you should have `R` installed on your machine. Preferably alongside `RStudio`. It is recommended that you use versions of `R` \> 4.0.0. So if you haven't updated recently, it might be a good time to do so.

If you are using a DEECA machine and and you do not have an open workstation then you may find it difficult to install `R`, `RStudio` and a range of `R` packages. It is recommended you have a workstation setting that allows you to download and install programs (like `R` and `R` packages). This is something you need to apply for through the DEECA IT service portal if you have not already. You should also have `Rtools` installed. Below we list a step-by-step guide to the installation of system dependencies for Windows if they are not already installed:

1.  Install `R` (preferably version \> 4.0.0): <https://cran.rstudio.com/>\
2.  Download the version of `Rtools` matching your `R` installation: <https://cran.r-project.org/bin/windows/Rtools/>\
3.  Install `RStudio`: <https://posit.co/download/rstudio-desktop/>

# Package Dependencies

For our workshop we will probably be downloading some packages on the fly, but if you want to make sure you have the relevant packages installed then you can run the following chunk:

```{r  pkg, eval = FALSE}
install.packages(pkgs = c("tidyverse",
                          "broom",
                          "mapview", 
                          "shiny", 
                          "rmarkdown")
))
```


# Workshop prerequisites




# Ummm, pretend that I forgot everything about Rstudio (asking for a mate)...

introduction to Rstudio use...\<\>

<br>

# ...and maybe a refresher on R as well??

introduction to Rstudio use...\<\>

<br>

# Wait, what is this Posit thing? I thought it was Rstudio

switch from Rstudio to Posit...<https://posit.co/>

Rmarkdown to Quarto

Rstudio IDE to Positron <https://github.com/posit-dev/positron/wiki>

<br>

# For f\*sake, now what is this Github thing?

couple notes on navigating Github...<https://github.com/bfanson>

# My brain hurts...

Zen out while checking out the 367 Collins Peregrines....

<iframe width="657" height="370" src="https://www.youtube.com/embed/VLLNNTjaSEA" title="367 Collins Falcons 2024 - South facing" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>

<br>

Panama bird/bat feeders....

<iframe width="654" height="460" src="https://www.youtube.com/embed/VfFfS64rtZE" title="Panama Fruit Feeder Cam at Canopy Lodge powered by EXPLORE.org" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>


<br>


