---
title: "Additional info on NA and NaN"
description: |
output:
  distill::distill_article:
    self_contained: true
    toc: true # table of content true
    toc_float: true # make 
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    code_folding: false # awesome buttons to show/hide the code
draft: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Advanced topic(s)

### Importing lots of files with same (or similar-ish) structure

There are times that you will want to import 10s or 100s of files that are structurally the same (e.g. camera data, acoustic tagging data). I use the `purrr` package on almost daily basis. This package was created to help simplify the `apply` family of functions (e.g. `apply`,`lapply`,`sapply`,`mapply`) into a more intuitive nomenclature (like the rest of tidyverse). One website that gives a tutorial is <https://jennybc.github.io/purrr-tutorial/> (but several others out there).

```{r import_multi, eval=F}

  v_list <- dir('data/raw/tagging',full.names = T)   # get a list of files to import.  use full.names=T to relative path
  importData <- function(x) read_csv( x ) %>% mutate(file_id=x) # create the function that imports a single file
  ds_all <- purrr::map_df( v_list, ~importData(.x) )   # loop through each filename in v_list, combine together into new dataset

```

Note - you could do this using for-loops but it takes more code.

### Excel files and importing color metadata

First off, **just don't use cell styles (color)** in excel as a way to store data. For instance, using color to indicate treatment.

That being said, you can extract that information using `tidyxl` package. I have had good luck with


