---
title: "Additional info on NA and NaN"
description: |
output:
  distill::distill_article:
    self_contained: true
    toc: true # table of content true
    toc_float: true # make 
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    code_folding: false # awesome buttons to show/hide the code
draft: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Historical context

-   base R [the packages included when you install R the first time]
-   `base`, `stats`, `grid`, `utils`, ...

```{r base_r, eval=F}
# example of getting a plot using base
  ds <- mtcars  # grab the built-in mtcars dataset
  ds <- ds[, c('mpg','hp')]   # select just the columns you want
  ds$l_100km <- 1/( ds$mpg/3.78 * 1.6 )*100 # convert mile per gallon to liters / 100km
  plot( ds$l_100km ~ ds$hp )  # plot relationship between horsepower and km/l

```

-   quirky things about R: factors vs character, NAs,

-   lubridate fixed date quirks

-   lack of consistency within base and between packages

-   data argument might be first or last

## Basic principles

-   human-centered

    -   readable to humans

    -   uses verbs like `mutate`, `select`, `filter`, `summarise`

    -   uses sentence like structure: noun -\> verb -\> verb [ Paul -\> walks (to work) -\> sets up ( his computer) -\> creates a plot -\> saves plot (to a file) -\> calls it day -\> walks (home) ]

-   consistent

    -   when using tidy functions, you get consistent output data structure (called a tibble)

```{r  eval=F}
  # an example of tidyverse consistent structure
  class(mtcars)   # data.frame
  head(mtcars)    # notice the rowlabels
  View(mtcars)    # see the rownames in the spreadsheet
  as_tibble(mtcars,rownames = 'make')  # tidyverse datasets (tibbles) never have rownames

```

-   

-   composable (chunks)

## Tibbles

-   I mentioned briefly above about "tibbles"

    tibble is data.frame v2.0

    -   better printing of the data
    -   handles some of the quirky things that caught up people
    -   column names

## Pipes: %\>% (magittr package)

A key component of tidyverse world is the concept of "pipes". The pipe connects multiples steps into a single sequence of commands.

<aside>Just to make things harder, a new pipe `|>` has been introduced. It is very similar to `%>%` but it does act a little different. We will just teach `%>%` because we are set in our ways</aside>

### Building "sentences"

```{r, eval=F}
  library(tidyverse)
  head(mtcars)
  mtcars %>% head(10)  # get mtcars, pass to head function, take the first 10 rows
  
# sentence using tidy functions we will dive into below...just follow basic logic
  mtcars %>% 
    slice_head(n=10) %>%  # take first 10 rows   
    mutate( mpl = round( mpg/3.79,2 ) ) %>%  # convert mpg to miles per liter (mpl)
    select( mpg, mpl )  # keep just the mpg and mpl columns 

# placeholder "."
  mtcars %>% head(5)    
  mtcars %>% head(.,5)  # same as above but explicitly using . to denote the input

  # example where you want to input the dataset not into first argument spot
  mtcars %>% lm(mpg~hp) %>% summary()  # fails as it is trying lm(formula=mtcars, data=mpg~hp )
  mtcars %>% lm(mpg~hp, data=.) %>% summary()  # referencing the input by "."

```

### "Odd" looking uses

With `%>%` it is worth being aware that you will see different variations that look odd. I show a few examples below.

<aside>Note that you can have different variations using `%>%` that might look odd. E.g. `data %>% head` or `data %>% .$var`. Just be beware</aside>

```{r eval=F}

  mtcars %>% head  # lack of brackets
  mtcars %>% head(n=3) %>%  bind_rows( ., .)   # multiple "." example
  mtcars %>% .$mpg   # grabs a specific column - combining tidyverse with base approaches 
  mtcars %>% {c( mean=mean(.$mpg), median=median(.$mpg) ) }  # curly bracket example using "." and $ - prevents putting . into c(., mean(.$mpg), median(.$mpg))


```

### old vs. new: %\>% vs. \|\>

\|\> is considered to be "simple" now compared to %\>%

are some differences mainly associated with referring to the passed

Tip - set a shortcut for... Ctrl-Shift-M

```{r , eval=F}
  library(tidyverse)
  head(mtcars)
  mtcars %>% head()
  mtcars |> head() # newer version 

# examples of differences (advanced)
  mtcars %>% head # works
  mtcars |> head  # fails  (must have brackets)
  mtcars |> head()  # works  (must have brackets)

  #subselecting
  mtcars %>% .$mpg
  mtcars |>  .$mpg  # $ not supported
  mtcars |>  pull(mpg)  # use pull

  mtcars %>% head(x=., 10) # . works for %>% 
  mtcars |>  head(x=., 10)  # . placeholder does not work
  mtcars |>  head(x=_, 10)  # _ placeholder instead
  mtcars |>  head(n=10, x=_)  # _ placeholder instead
  mtcars  %>%   head(n=10, .)  # can place without argument specified
  mtcars  %>%   head(n=10, _)  # must you argument if not in first position

    
```
