---
title: "Dataset"
description: |
  Describes the dataset for the workshop
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

# Project data

If you want to start from scratch, click on link below to download project data files and go from there. _Each day we will provide a Rstudio project that will have the data as well_


```{r downloaddata, echo=FALSE, code_folding = FALSE}
library(downloadthis)
  download_link(
    link='https://delwpvicgovau-my.sharepoint.com/:u:/g/personal/ben_fanson_deeca_vic_gov_au/ET-SqpFmklhNqbUxXAwmHcMBWR2OtEDry-OVXha4fsJ5dg?e=cVKgEY',
    output_name = "Attendance list.xlsx",
    button_label = "Download project data",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

<br>

# Project overview
```{r echo=F, fig.height=4, fig.width=4}
  sf_site <- sf::read_sf('data/sf_site.gpkg')
  sf_cma <- VicmapR::vicmap_query(layer='open-data-platform:cma100') %>% collect()


```

## Purpose

As project


## Experimental design

```{r echo=F}
  n_wetlands <- n_distinct(sf_site$wetland)
  n_crash <- sum(sf_site$grazing=='Crash')
  n_graze <- sum(sf_site$grazing=='Crash')
  
```
Across western and eastern Victoria, a total of `r n_wetlands` wetlands were surveyed.  At each wetland




```{r echo=F, fig.height=4, fig.width=7}
  #sf_cma <- VicmapR::vicmap_query(layer='open-data-platform:cma100') %>% collect()
  #sf::write_sf(sf_cma, 'sf_cma.gpkg')
  mapview::mapview(sf_cma, zcol='cmaname', layer.name='cma') + 
    mapview::mapview(sf_site, zcol='grazing', layer.name='site' )

```


# Data overview

## Main Excel file


## CSV survey


## CSV 


```{r}

# kableExtra::scroll_box( kableExtra::kable(cars[1:10,],'html'),
#                         height="200px", width='180px')


```




