---
title: "Day 4: Shiny apps"
description: |
  A short description of the post.
author:
  - name: Ben Fanson
    url: https://bfanson.github.io/2024DADAworkshop/
    affiliation: Arthur Rylah Institute
    affiliation_url: https://www.ari.vic.gov.au/
date: 2024-09-01
output:
  distill::distill_article:
    self_contained: true
    toc: true # table of content true
    toc_float: true # make 
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    code_folding: true # awesome buttons to show/hide the code
draft: false
slug: fanson2024dada
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, collapse = TRUE)
```

# General thoughts

1)  For Shiny Apps, the code looks foreign and a bit are you serious after all this time I spent learning R coding

2)  Once you get the basic structure, you can easily get started and then it is "Trial and Error"

# Day Objectives

1)  Understand the core structure of Shiny apps: UI, Server, Reactivity

2)  Learn how to create and run an App locally

3)  Learn how to deploy your App on ARI's shiny account


# Think about how shiny Apps might be useful

## Visualising data for clients 

Ibis Tracker - Nev masterpiece

https://arisci.shinyapps.io/ibisTracker/

## Data sharing - upload 



## Model simulations


# Reminder on a few R fundamentals

LISTS

# Creating a shiny app using Rstudio

```{r, class.source="bg-success"}
#install.packages("shiny")

```

You can find a variety of cheatsheets at <https://posit.co/resources/cheatsheets/>

```{r cheatsheet, out.width='100%', out.height='200%'}
  knitr::include_graphics("shiny.pdf")
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
s


```{r}
# library(shiny) 
# ui <- fluidPage(   "Hello, world!" ) 
# server <- function(input, output, session) { } 
# shinyApp(ui, server)
```

# steps

1)  add UI controls

- this is separated by commas!!!!
- So many nested brackets!!!!

2)  add "behaviour" Shiny uses reactive programming to make apps interactive.

- not separated by commas...more like R programming but order does not matter!!!

3)  

you put R code in renderFunction

# Shiny Templates

-   how might you look it up

extras at \< <https://github.com/nanxstats/awesome-shiny-extensions>\>

# Exercise

Let's create replicate Examples <https://shiny.posit.co/r/gallery/>

# Walk through of creating App for our dataset

## App design

## Let's build something


# Jumping back to workflow for Apps


This will become pretty useful for collaborations and your own self-learning. Also, we now have a ARI shiny account and consistency allows for increased sharing and understanding.


# Useful interactive packages for shiny

## leaflet

## plotly

```{r}
  library(ggplot2)
  library(plotly)
  f <- ggplot( cars, aes(speed, dist)  ) + geom_point() + geom_smooth()
   plotly::ggplotly(f)

```


# Deploy your App

## ARI shiny account

## rsconnect package

```{r eval=F, class.source="bg-danger", class.output="bg-warning"}
install.packages('rsconnect')
```

``` markdown
install.packages('rsconnect')
```

``` markdown
rsconnect::setAccountInfo(name="<ACCOUNT>", token="<TOKEN>", secret="<SECRET>")
```

# Shot bit on layouts

https://shiny.posit.co/r/layouts/

# Beyond showing results

## Uploading files

## Saving user inputs

## Downloading

# Additional resources

For useful resources: <https://shiny.posit.co/r/articles/> For useful examples: <https://shiny.posit.co/r/articles/>

see <https://mastering-shiny.org/>

<aside>Asides can be useful for preso.</aside>
