---
title: "Day 4: Shiny apps"
description: |
  A short description of the post.
author:
  - name: Ben Fanson
    url: https://bfanson.github.io/2024DADAworkshop/
    affiliation: Arthur Rylah Institute
    affiliation_url: https://www.ari.vic.gov.au/
date: 2024-09-01
output:
  distill::distill_article:
    self_contained: true
    toc: true # table of content true
    toc_float: true # make 
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    code_folding: true # awesome buttons to show/hide the code
draft: false
slug: fanson2024dada
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, collapse = TRUE)
```

# General thoughts

1) For Shiny Apps, the code looks foreign and a bit are you serious after all this time I spent learning R coding

2) Once you get the basic structure, you can easily get started and then it is "Trial and Error"

# Day Objectives

1) Understand the core structure of Shiny apps: UI, Server, Reactivity

2) Learn how to create and run an App locally

3) Learn how to deploy your App on ARI's shiny account




# Reminder on a few R fundamentals

LISTS 

# Creating a shiny app using Rstudio



# Header1

```{r, class.source="bg-success"}
#install.packages("shiny")

```

You can find a variety of cheatsheets at <https://posit.co/resources/cheatsheets/>

```{r cheatsheet, out.width='100%', out.height='200%'}
  knitr::include_graphics("shiny.pdf")
```

```{r}        
# library(shiny) 
# ui <- fluidPage(   "Hello, world!" ) 
# server <- function(input, output, session) { } 
# shinyApp(ui, server)
```


# steps
1) add UI controls
2) add "behaviour"
Shiny uses reactive programming to make apps interactive. 
3)

you put R code in renderFunction


# Shiny Templates

- how might you look it up

extras at < https://github.com/nanxstats/awesome-shiny-extensions>

# Exercise

Let's create replicate 
Examples <https://shiny.posit.co/r/gallery/>



# Additional resources


see <https://mastering-shiny.org/>

<aside>Asides can be useful for preso.</aside>
